Procedural Terrain \\ {\large Final project for class CS6491 Computer Graphics}

### 2015-11-30 Paper/Paper.tex
modified the constant values to match the new ones in the implementation.

\subsubsection{Erosion, deposition and sediment transportation}\label{ErosionDeposition}
To compute the sediment erosion and deposition, we first define the slope at a cell as:
\begin{equation}
\begin{array}{rl}
	\alpha[x,y] = \max(&|b[x,y]-b[x-1,y]|,\\
	                   &|b[x,y]-b[x+1,y]|,\\
										 &|b[x,y]-b[x,y-1]|,\\
										 &|b[x,y]-b[x,y+1]|\ )
\end{array}
\label{eq:ErosionSediment1}
\end{equation}
Then the sediment transportation capacity $C$ is calculated as:
\begin{equation}
	C[x,y]=K_c \cdot \alpha[x,y] \dot |\textbf{v}[x,y]|
\label{eq:ErosionSediment2}
\end{equation}
In our experiments, we set the sediment capacity constant $K_c$ to 0.5.
We recommend to clamp the values of $\alpha$ to be between e.g. 0.005 and 0.1 before computing $C$. By this, the erosion on steep areas does not grow to infinity and on flat areas some erosion still takes place.

Then we compare $C$ to the suspended sediment $s_t$.
For $C>s_t$, sediment is eroded using:
\begin{equation}
\begin{array}{rcl}
	b_{t+\Delta t} &=& b_t - K_s (C-s_t) \\
	s' &=& s_t + K_s (C-s_t)
\end{array}
\label{eq:ErosionSediment3_1}
\end{equation}
And for $C<s_t$, sediment is dissolved:
\begin{equation}
\begin{array}{rcl}
	b_{t+\Delta t} &=& b_t + K_d (s_t-C) \\
	s' &=& s_t - K_d (s_t-C)
\end{array}
\label{eq:ErosionSediment3_2}
\end{equation}
$K_s$ and $K_d$ are erosion and deposition constants. In our experiments, we use $K_s=K_d=0.1$.

@@

\subsubsection{Evaporation}
As a last step, some water is evaporated in the air every step. We model this simply as
\begin{equation}
	d_{t+\Delta t}[x,y] = d_{t+\Delta t}[x,y] \cdot (1-K_e \cdot \Delta t)
\label{eq:ErosionEvaporation}
\end{equation}
With $K_e$ being an evaporation constant, e.g. $K_e=0.005$.

@@

\subsubsection{non-uniform erosion and deposition constants}
In \ref{ErosionDeposition} we assumed the erosion constant $K_s$ and the deposition constant $K_d$ to be constant.
However, in a natural soil, this is not the case.
\begin{center}
	\includegraphics[width=0.3\textwidth]{images/Soil_profile}
	\footnote{https://upload.wikimedia.org/wikipedia/commons/9/95/Soil\_profile.png}
\end{center}
The top layers (soil) are eroded more easily than the bottom layers (stone). Newly dissolved sediment is eroded even better again because it consists mostly of loose stones and sand.
We model this by representing $K_s$ and $K_d$ as functions of the amount of eroded terrain $\Delta b$.
\begin{equation}
\begin{array}{rcl}
	\Delta b_t[x,y] &=& b_\text{original}[x,y] - b_t[x,y] \\
	K_s(\Delta b) &=& K_{s,\text{base}} \cdot 2^{-\Delta b / f_s} \\
	K_d(\Delta b) &=& K_{d,\text{base}} \cdot 2^{\Delta b / f_d} \\
\end{array}
\label{eq:KsKd}
\end{equation}
$K_{s,\text{base}}$ and $K_{d,\text{base}}$ are set to 0.1 as before; $f_s$ and $f_d$ specifies the amount of terrain that has to be eroded/deposited before the speed of the erosion/deposition is halved.

>>>>

\subsubsection{Erosion, deposition and sediment transportation}\label{ErosionDeposition}
To compute the sediment erosion and deposition, we first define the slope at a cell as:
\begin{equation}
\begin{array}{rl}
	\alpha[x,y] = \max(&|b[x,y]-b[x-1,y]|,\\
	                   &|b[x,y]-b[x+1,y]|,\\
										 &|b[x,y]-b[x,y-1]|,\\
										 &|b[x,y]-b[x,y+1]|\ )
\end{array}
\label{eq:ErosionSediment1}
\end{equation}
Then the sediment transportation capacity $C$ is calculated as:
\begin{equation}
	C[x,y]=K_c \cdot \alpha[x,y] \cdot |\textbf{v}[x,y]|
\label{eq:ErosionSediment2}
\end{equation}
In our experiments, we set the sediment capacity constant $K_c$ to 0.1.
We recommend to clamp the values of $\alpha$ to be between e.g. 0.005 and 0.5 before computing $C$. By this, the erosion on steep areas does not grow to infinity and on flat areas some erosion still takes place.

Then we compare $C$ to the suspended sediment $s_t$.
For $C>s_t$, sediment is eroded using:
\begin{equation}
\begin{array}{rcl}
	b_{t+\Delta t} &=& b_t - K_s (C-s_t) \\
	s' &=& s_t + K_s (C-s_t)
\end{array}
\label{eq:ErosionSediment3_1}
\end{equation}
And for $C<s_t$, sediment is dissolved:
\begin{equation}
\begin{array}{rcl}
	b_{t+\Delta t} &=& b_t + K_d (s_t-C) \\
	s' &=& s_t - K_d (s_t-C)
\end{array}
\label{eq:ErosionSediment3_2}
\end{equation}
$K_s$ and $K_d$ are erosion and deposition constants. In our experiments, we use $K_s=K_d=0.002$.

@@

\subsubsection{Evaporation}
As a last step, some water is evaporated in the air every step. We model this simply as
\begin{equation}
	d_{t+\Delta t}[x,y] = d_{t+\Delta t}[x,y] \cdot (1-K_e \cdot \Delta t)
\label{eq:ErosionEvaporation}
\end{equation}
With $K_e$ being an evaporation constant, e.g. $K_e=0.1$.

@@

\subsubsection{non-uniform erosion and deposition constants}
In \ref{ErosionDeposition} we assumed the erosion constant $K_s$ and the deposition constant $K_d$ to be constant.
However, in a natural soil, this is not the case.
\begin{center}
	\includegraphics[width=0.3\textwidth]{images/Soil_profile}
	\footnote{https://upload.wikimedia.org/wikipedia/commons/9/95/Soil\_profile.png}
\end{center}
The top layers (soil) are eroded more easily than the bottom layers (stone). Newly dissolved sediment is eroded even better again because it consists mostly of loose stones and sand.
We model this by representing $K_s$ and $K_d$ as functions of the amount of eroded terrain $\Delta b$.
\begin{equation}
\begin{array}{rcl}
	\Delta b_t[x,y] &=& b_\text{original}[x,y] - b_t[x,y] \\
	K_s(\Delta b) &=& K_{s,\text{base}} \cdot 2^{-\Delta b / f_s} \\
	K_d(\Delta b) &=& K_{d,\text{base}} \cdot 2^{\Delta b / f_d} \\
\end{array}
\label{eq:KsKd}
\end{equation}
$K_{s,\text{base}}$ and $K_{d,\text{base}}$ are set to 0.002 as before; $f_s$ and $f_d$ specifies the amount of terrain that has to be eroded/deposited before the speed of the erosion/deposition is halved.
